Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting

### 2025-04-05 products/manuscript/Manuscript.qmd
Le Projet 2

Efforts to mitigate the impact of malaria in pregnancy have centered on preventive strategies such as the use of insecticide-treated bed nets (ITNs) and chemoprophylaxis through intermittent preventive treatment during pregnancy (IPTp). Two widely used IPTp regimens are sulfadoxine-pyrimethamine (SP) and dihydroartemisinin-piperaquine (DP). While these preventive measures have shown significant benefits in reducing the risk of malaria-related complications, the effectiveness of these regimens may not be uniform across all contexts. In particular, emerging evidence suggests that the choice of IPTp regimen may influence how malaria episodes affect pregnancy outcomes. However, this potential effect modification remains underexplored in current literature. Additionally, maternal characteristics such as gravidity—the number of times a woman has been pregnant—may also play a role in shaping birth outcomes. Prior research has suggested that previous pregnancy experience may offer protective benefits against adverse outcomes, possibly due to improved physiological adaptation or better health-seeking behavior.

@@

*Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes*
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create the composite adverse outcome variable
promo_data_clean <- promo_data_clean %>%
  mutate(
    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
  )

@@

@@

**Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#library(broom)
library(dplyr)
library(knitr)
library(kableExtra)
library(here)

@@

**Table 4:  Regression Diagnostics for IPTp Treatment Effect**
```{r, echo=FALSE, message=FALSE, warning=FALSE}

@@

**Table 5: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/Images/Table- Comparing Models .png")

@@

**Table 4: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**

@@

*Table 6: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)*
```{r, echo=FALSE, message=FALSE, warning=FALSE}

@@

*Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)*

>>>>

Efforts to mitigate the impact of malaria in pregnancy have centered on preventive strategies such as the use of insecticide-treated bed nets (ITNs) and chemoprophylaxis through intermittent preventive treatment during pregnancy (IPTp). Two widely used IPTp regimens are sulfadoxine-pyrimethamine (SP) and dihydroartemisinin-piperaquine (DP). While these preventive measures have shown significant benefits in reducing the risk of malaria-related complications, the effectiveness of these regimens may not be uniform across all contexts. In particular, emerging evidence suggests that the choice of IPTp regimen may influence how malaria episodes affect pregnancy outcomes. However, this potential effect modification remains underexplored in current literature. Additionally, maternal characteristics such as gravidity, the number of times a woman has been pregnant may also play a role in shaping birth outcomes. Prior research has suggested that previous pregnancy experience may offer protective benefits against adverse outcomes, possibly due to improved physiological adaptation or better health-seeking behavior.

@@

**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create the composite adverse outcome variable
promo_data_clean <- promo_data_clean %>%
  mutate(
    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
  )

@@

**Table 3: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
```{r,  echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(here)

@@

**Table 4: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#library(broom)
library(dplyr)
library(knitr)
library(kableExtra)
library(here)

@@

**Table 5:  Regression Diagnostics for IPTp Treatment Effect**
```{r, echo=FALSE, message=FALSE, warning=FALSE}

@@

**Table 6: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/Images/Table- Comparing Models .png")

@@

**Table 7: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**

@@

**Table 8: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
```{r, echo=FALSE, message=FALSE, warning=FALSE}

@@

**Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
